---
title: "Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Here we give an introduction to the package multiple testing in R! We do a sample
analysis and look at trends in gridded temperature data, provided by NASA here:

https://data.giss.nasa.gov/pub/gistemp/GHCNv3/gistemp1200_ERSSTv5.nc.gz

The data in the package has been preprocessed so that we have yearly averages 
in a 3d array with dimensions (lon, lat, t), where t is time. We start by 
taking a quick look:


```{r setup}
library(PerMuTe)
str(temp_gistemp)
```

```{r setup}
image(temp_gistemp[,,1])
```
Our objective is to determine **where** there are significant increasing or 
decreasing trends. For this, we perform a trend test at each grid cell. **Note** 
when doing this, we are repeating a test thousands of times and the existance of 
false positives is **guaranteed**. Ideally, we want to control the probability 
that our results have false positives. This probability is what we control, at 
the desired $\alpha$, when we apply a correction for multiple testing. 

The popular Bonferroni correction and related methods are **too conservative** 
when performing thousands of tests, so we follow two novel permutation methods, 
as described in

> Cort\'es, J., Mahecha, M., Reichstein, M. et al. Accounting for multiple 
testing in the analysis of spatio-temporal environmental data. Environ Ecol Stat 27, 
293–318 (2020). https://doi.org/10.1007/s10651-020-00446-4

These new methods generally perform better. Note that the temperature example 
is also shown in the paper above. 


## Data and Method 
* Gridded temperature data for the world, averaged yearly from 1951-2018
* Mann-Kendall trend test at each grid cell, at $\alpha_{local} = 0.05$  
* Permutation correction for multiple testing, at $\alpha_{global} = 0.05$: 
  + Based on the maximum statistic (maxT)
  + Based on the supra-threshold cluster size (STCS)
  
Note: there are two $alpha$'s above. The $\alpha_{local}$ controls the **individual**
test's significance, while the $\alpha_{global}$ controls the **overall** 
probability of having a false positive. I recommend that these two be equal to 
avoid any confusion in the interpretation of the results. 

The idea is to derive the distribution of maxT and STCS via permutations. With 
these distributions we can establish thresholds to control the overall false 
positives. 

For each permutation method we do the following:

maxT           | STCS
-------------- | --------------
Permute images |	Permute images
Calculate test statistic at each grid cell | Calculate test statistic at each grid cell
Keep the maximum statistic among all grid cells |	Keep the size of the largest cluster of significant grid cells
Repeat **N** times |	Repeat **N** times

Where **N** is typically 1000. Because we are essentially repeating the analysis
**N** times, this can be time consuming. For the purposes of this tutorial, we will 
set **N** to be 100. 

The above is steps are what is already implemented for you in the package.

## R setup

There are several inputs to the main function of the r package: 
`multiple_testing correction`. We go over each of them briefly. 

data 
:The input data must be a 3d array, where the dimensions are (lon, lat, t), in 
that order. 

fx
:This is the desired function to be applied at each grid cell. You need to specify 
your own function in R. This requires a little more programming but allows for 
any type of analysis at the grid cell level. Permutation methods are 
computationally expensive, so we try to keep the function to its minimum 
requirements. A sample function to perform the MK trend test with an adjustment
for serial autocorrelation is included: 

```{r}
sample_mk_function
```
method
:Which correction for multiple testing should be applied. Defaults to all. 
Options are: c("maxT", "stcs", "bonferroni", "bh", "by", "holmes", "hochberg")

nperm
:number of permutations to perform. Default to 1000, which is a typical value. 
For testing the function a lower value is better, e.g., 10. 

alpha_local
:significance level at each grid cell. 

alpha_global
:significance level for the overall study area

null_distribution
:for bonferroni and related methods, a distribution is needed to derive the 
p-values of the test statistics. Supported options are "normal" and "t". For 
example, the Mann-Kendall's S can be converted to a Z score, and so the 
distribution is "normal". 

seed
:set a seed to get same results every time. 

block_size
:length of block for performing block permutations. 

verbose
:a counter that prints when it is done with every 10 permutations. Defaults to 
TRUE
                                   

## Running the analysis

Now that we have explained everything, lets run the analysis in R!

```{r}
results<- multiple_testing_correction(data = temp_gistemp, 
                                      fx = sample_mk_function, 
                                      method = "all", 
                                      nperm = 10, 
                                      alpha_local = 0.10, 
                                      alpha_global = 0.10, 
                                      null_distribution = "normal")
```
When verbose=TRUE, the function prints when it has completed multiples of 10 permutations

In the results we have a named list of 2d matrices matching the lon, lat dimensions of the 
input data. Each matrix corresponds to each specified method, where each entry
is TRUE or FALSE depending on whether or not it is significant. The first two matrices 
are the test statistic at each grid cell and their corresponding p-values. 

The End!


## Plotting results

To look at our data results we can do 

```{r}
image(results$test_statistic)
image(results$maxT)
image(results$stcs)
```

Note that Bonferroni and other methods detect much less, for example: 

```{r}
image(results$bonf)
```

Now, we could continue the analysis with the selected grid cells, for example, 
by estimating the slope at each of the selected grid cells and summarizing by 
continent. 

We end this intro with a nicer map:

```{r}
#library(raster)
library(ggplot2)
library(sp)
library(sf)
#library(rgeos)
library(maps)
library(reshape2)
# world <- ne_countries(scale = "medium", returnclass = "sf")
temperature_results_stcs<- melt(results$test_statistic)
spPoints<- SpatialPointsDataFrame(coords = data.frame(temperature_results_stcs[,1:2]),
                                  data = data.frame(mk_z=temperature_results_stcs$value),
                                  proj4string = CRS("+init=epsg:4326"))
polys<- as(SpatialPixelsDataFrame(spPoints,spPoints@data, tolerance = 0.149842),"SpatialPolygonsDataFrame")
polys_sf = as(polys, "sf")

library(maptools)
borders  <- map("world", fill = T, plot = F)
IDs      <- seq(1,1627,1)
borders  <- map2SpatialPolygons(borders, IDs=borders$names, 
                               proj4string=CRS("+proj=longlat +datum=WGS84")) %>% 
            as("sf")
my_theme <- theme_bw() + theme(panel.ontop=TRUE, panel.background=element_blank())

ggplot(polys_sf) +
  geom_sf(aes(fill = mk_z), color = "transparent") + 
  geom_sf(data = borders, fill = "transparent", color = "black") +
  scale_fill_distiller(palette='Spectral') +
  ggtitle("Mann-Kendall's Z statistic") +
  #borders('world', colour='black')+
  coord_sf(crs = st_crs(4326), xlim = c(-180, 180), ylim = c(-90, 90))+
  my_theme 


  

```

The above map was generated following

https://stackoverflow.com/questions/43612903/how-to-properly-plot-projected-gridded-data-in-ggplot2

The End!
